---
title: "Preliminary-results.Rmd"
author: "Group_4"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary Results

1. Title

2. Team members names

3. Introduction. Explain the origin and structure of your data.

4. Objectives and hypotheses.

5. Methods. What have you done? And what did you use to do it? Step by step, so anyone could do it again by reading this part.

**Objectives**

1. Evaluate the effects of the water stress treatments on the growth, development and photosynthetic activity in nine different plant species.
2. Compare the responses to drought conditions between cultivated crops and wild weeds, and determine if there are significant differences.

**Hypotheses**

a. H0.1: There will be no significant effects of the water stress treatments on the growth, development and photosynthetic activity in the nine tested plant species. 
b. H0.2: There will be no significant differences on the growth, development and photosynthetic activity of plants from the water treatments between cultivated and wild species. 
c. H1.1: There will be significant effects of the water stress treatments on the growth, development and photosynthetic activity in the nine tested plant species. 
d. H1.2: There will be significant differences in growth, development and photosynthetic activity of plants from the water treatments between cultivated and wild species.

**Intended analyses**

#### a. H0.1: There will be no significant effects of the water stress treatments on the growth, development and photosynthetic activity in the ten tested plant species.

 1. Investigate the effects of the 3 water stress treatments 
 
 Plant Parameters to be used:
 
 - Leaf number 
 - Chlorophyll content (W6)
 - Root length
 - Aerial weight (fresh and dry)
 - Root weight (fresh and dry)
 - Water content
 
 Categorical parameters:
 - Species
 - Date
 - Week 
 - Use
 - Treatment
 -Plant ID
 
  a. Modify our data table so it includes the parameters described before.
  
  b. Carry out the actual statistical tests:

H1.0:

    - One way ANOVA 
    Categorical variables: treatments
    Dependent variables: plant parameters
    Calculate one parameter at a time 
    
    parameters to control:
    - Date
    - Species
    
    - Shapiro–Wilk test for normality of residuals 
    - TukeyHSD test (PostHoc test) to recognize different groups from ANOVA
    - Kruskal Wallis as Non-parametric test

H2.0:

    - Two way ANOVA 
    Categorical variables: treatment and use
    Dependent variables: plant parameters
    Calculate one parameter at a time 
    
    parameters to control:
    - Date
    - Species
    
    - Shapiro–Wilk test for normality of residuals 
    - TukeyHSD test (PostHoc test) to recognize different groups from ANOVA
    - Kruskal Wallis as Non-parametric test    
    
6. Results. Figures and tables with captions and description of what do they mean.

7. Discussion. From your objectives.

8. Conclusions.

```{r cars}
###
```

## Plots


